Robust calibration of numerical models based on relative regret

### 2019-11-22 article.tex
modif name .tex, update .gitignore, del pdf

KDE is a non-parametric estimation technique based on the use of a kernel function $f$. Assuming an isotropic kernel, the estimation has the following form:
\begin{equation}
\hat{p}_{\mathbf{K}^*}(\mathbf{k}) = \frac{1}{Nh^{\dim \Kspace}} \sum_{i=1}^N f\left(\frac{\mathbf{k} - \mathbf{k}_{\mathbf{u}^i}}{h}\right)
\end{equation}
where $h$ is the bandwidth. In a multidimensional setting, one usually consider a kernel based on the product of 1D kernels, applied independently to all components: $f(\mathbf{k}) = \prod_{j=1}^{\dim \Kspace} f_{\mathrm{1D}} (k^{(j)})$
where $k^{(j)}$ is the $j$-th component of $\mathbf{k}$.  There is wide choice of available $f_{\mathrm{1D}}$, such as a rectangle function, $f_{\mathrm{1D}}(x) = 1 \text{ if } |x|< \frac{1}{2}$ and  $0$ elsewhere, or a Gaussian kernel $f_{\mathrm{1D}}(x)= \frac{1}{\sqrt{2\pi}}\exp(-x^2 / 2)$. One can notice that the rectangle function as a kernel yields non-smooth densities, very similar to the histogram of the minimisers.

@@

Compared to the plots for $J_{\mathrm{BH}}$ in Figure~\ref{fig:ratio_BH}, $\hat{\checka}_p$ exhibits a smoother behaviour for the $J_{\mathrm{BHswap}}$ function as no change of slope is easily discernable. The computation of the ratio advocates for a level of confidence $p_{\mathrm{maxratio}}$ close to $0.75$. The numerical values of the estimations $\hat{k}$ presented in Table~\ref{tab:recap_estimates_BHs} show that contrary to the $J_{\mathrm{BH}}$ function, the objective are concurrent.

@@

\begin{figure}[!h]
\centering
\includegraphics[width=.7\textwidth]{Figures/RobustSWEmis_plots.pdf}
\caption{Procedure of robust calibration for the shallow water problem. Top: contours of $J$, conditional minimisers and $\{(\mathbf{k},\mathbf{u}) | J(\mathbf{k},\mathbf{u}) = \hat{\alpha}_1 J^*(\mathbf{u})\}$. Middle: Conditional moments and histogram and KDE of the conditional minimisers. Bottom: $\hat{\Gamma}_{\alpha_p}$ for different levels $p$}
\label{fig:estimation_swe}
\end{figure}
\begin{table*}[!h]
\centering
\caption{Calibrated values of $\mathbf{k}$ according to different criteria}
\begin{tabular}{lr} \toprule
Estimator & Value \\ \midrule
$\hatkmpe$ & 0.249 \\
$\kest_{\mathrm{global}}$ & 0.290 \\
$\hat{\mathbf{k}}_{p},~p=1$ & 0.423 \\ 
$\hat{\mathbf{k}}_p,~p=0.90$ & 0.458 \\ 
$\hatkmean$ & 0.501 \\ 
$\hat{\mathbf{k}}_p,~p=0.80$ & 0.505 \\ 
$\hat{\mathbf{k}}_p,~p=0.70$ & 0.560 \\ 
$\hatkvar$ & 0.990\\ \bottomrule
\end{tabular}
\label{tab:recap_estimates_metaSWE}
\end{table*}
Similarly as for $\JBH$ and $\JBHS$, the profiles are depicted Figure~\ref{fig:profiles_swe}.
\begin{figure}[!h]
  \centering
  \includegraphics[width=.8\textwidth]{Figures/profile_swe}
  \caption{Profiles for the calibration problem}
\label{fig:profiles_swe}
\end{figure}
We can see that the performances of $\kest_1$ and $\hatkmean$ are very similar, but $\kest_1$ has better performances when $\mathbf{u} > 0.6$.
\clearpage
\subsection{Assessing the quality of the forecast of the calibrated model}
The calibration of the model has been performed using by integrating the model on a time-period $[0, T]$, called assimilation window. 
We now want to compare the quality of the different forecasts, from different calibrated bottom friction.
Those forecasts result from the integration of the numerical model between the time $T$ and a time $T_{\mathrm{pred}}$.

@@

\subsubsection{Squared forecast error}
Given two environmental conditions $\mathbf{u}$, used to run the computer simulation and $\mathbf{u}'$, used to generate the observations, the squared forecast error for the parameter $\mathbf{k}$ is
\begin{equation}
\label{eq:squared_forecast_error}
S_{\mathrm{pred}}(\mathbf{k},\mathbf{u},\mathbf{u}') = \left(\mathcal{M}_{\mathrm{pred}}(\mathbf{k},\mathbf{u}) - \mathcal{M}^{o}_{\mathrm{pred}}(\mathbf{u}')\right)^2
\end{equation}
By averaging over $\mathbf{u}$ and $\mathbf{u}'$ using a Monte-Carlo approximation, we can define the mean squared forecast error, defined on every point of the spatial domain, and at every time-steps.
\begin{equation}
  \label{eq:squared_forecast_error_averaged}
  S(\mathbf{k}) = \frac{1}{N_{\mathbf{u}} N_{\mathbf{u}'}}\sum_{\mathbf{u},\mathbf{u}'\sim \mathbf{U}} S_{\mathrm{pred}}(\mathbf{k},\mathbf{u},\mathbf{u}') = \frac{1}{N_{\mathbf{u}} N_{\mathbf{u}'}}\sum_{\mathbf{u},\mathbf{u}'\sim \mathbf{U}} \left(\mathcal{M}_{\mathrm{pred}}(\mathbf{k},\mathbf{u}) - \mathcal{M}^{o}_{\mathrm{pred}}(\mathbf{u}')\right)^2
\end{equation}
The mean squared forecast error is then averaged over all time-steps between $T$ and $T_{\mathrm{pred}}$, indicating on which point of the domain the forecast error is greater as seen on the left plot of Figure~\ref{fig:forecast_squared_error}.
\begin{figure*}[!h]
\centering
\includegraphics[width=.9\textwidth]{Figures/forecast_error.pdf}
\caption{Quality of forecast, depending on the calibrated parameter. The left figures show the metric averaged over time, while the right figure show the mean forecast error, averaged over time and space}
\label{fig:forecast_squared_error}
\end{figure*}
We can see that the mean forecast error squared, on the right side, averaged over time and space is the smallest for $\kest_{\mathrm{global}}$, while $\kest_{1.0}$ performs slightly better $\hatkmean$.
\begin{figure}[!h]
  \centering
  \includegraphics[width=.9\textwidth]{Figures/forecast_CRPS}
  \caption{CRPS computed for different calibrated parameters}
\label{fig:forecast_crps}
\end{figure}

@@

\clearpage
\section*{Conclusion}  %% \conclusions[modified heading if necessary]
This paper deals with the problem of robust calibration of a computer code in the presence of uncertain inputs.

>>>>

KDE is a non-parametric estimation technique based on the use of a kernel function $f$. Assuming an isotropic kernel, the estimation has the following form:
\begin{equation}
\hat{p}_{\mathbf{K}^*}(\mathbf{k}) = \frac{1}{Nh^{\dim \Kspace}} \sum_{i=1}^N f\left(\frac{\mathbf{k} - \mathbf{k}_{\mathbf{u}^i}}{h}\right)
\end{equation}
where $h$ is the bandwidth. In a multidimensional setting, one usually consider a kernel based on the product of 1D kernels, applied independently to all components: $f(\mathbf{k}) = \prod_{j=1}^{\dim \Kspace} f_{\mathrm{1D}} (k^{(j)})$
where $k^{(j)}$ is the $j$-th component of $\mathbf{k}$.  There is wide choice of available $f_{\mathrm{1D}}$, and a popular choice is  %such as a rectangle function, $f_{\mathrm{1D}}(x) = 1 \text{ if } |x|< \frac{1}{2}$ and  $0$ elsewhere,
the Gaussian kernel $f_{\mathrm{1D}}(x)= \frac{1}{\sqrt{2\pi}}\exp(-x^2 / 2)$. % One can notice that the rectangle function as a kernel yields non-smooth densities, very similar to the histogram of the minimisers.

@@

Compared to the plots for $J_{\mathrm{BH}}$ in Figure~\ref{fig:ratio_BH}, $\hat{\checka}_p$ exhibits a smoother behaviour for the $J_{\mathrm{BHswap}}$ function as no change of slope is easily discernable. The computation of the ratio advocates for a level of confidence $p_{\mathrm{maxratio}}$ close to $0.75$. The numerical values of the estimations $\hat{k}$ presented in Table~\ref{tab:recap_estimates_BHs} show that, contrary to the $J_{\mathrm{BH}}$ function, the objective are concurrent.

@@

\begin{figure}[!h]
\centering
\includegraphics[width=.7\textwidth]{Figures/RobustSWEmis_plots.pdf}
\caption{Procedure of robust calibration for the shallow water problem. Top: contours of $J$, conditional minimisers and $\{(\mathbf{k},\mathbf{u}) | J(\mathbf{k},\mathbf{u}) = \hat{\alpha}_1 J^*(\mathbf{u})\}$. Middle: Conditional moments and histogram and KDE of the conditional minimisers. Bottom: $\hat{\Gamma}_{\alpha_p}$ for different levels $p$}
\label{fig:estimation_swe}
\end{figure}
\begin{table*}[!h]
\centering
\caption{Calibrated values of $\mathbf{k}$ according to different criteria}
\begin{tabular}{lr} \toprule
Estimator & Value \\ \midrule
$\hatkmpe$ & 0.249 \\
$\kest_{\mathrm{global}}$ & 0.290 \\
$\hat{\mathbf{k}}_{p},~p=1$ & 0.423 \\ 
$\hat{\mathbf{k}}_p,~p=0.90$ & 0.458 \\ 
$\hatkmean$ & 0.501 \\ 
$\hat{\mathbf{k}}_p,~p=0.80$ & 0.505 \\ 
$\hat{\mathbf{k}}_p,~p=0.70$ & 0.560 \\ 
$\hatkvar$ & 0.990\\ \bottomrule
\end{tabular}
\label{tab:recap_estimates_metaSWE}
\end{table*}
Similarly as for $\JBH$ and $\JBHS$, the profiles are depicted Figure~\ref{fig:profiles_swe}.
\begin{figure}[!h]
  \centering
  \includegraphics[width=.8\textwidth]{Figures/profile_swe}
  \caption{Profiles for the calibration problem}
\label{fig:profiles_swe}
\end{figure}
We can see that the performances of $\kest_1$ and $\hatkmean$ are very similar, but $\kest_1$ has better performances when $\mathbf{u} > 0.6$. When comparing with the global optimiser $\kest_{\mathrm{global}}$, $\kest_1$ performs better when $\mathbf{u}<0.625$. As $\mathbf{U}$ is uniformly distributed on $\Uspace=[0.5,0.7]$, we can graphically estimate that the probability that $\kest_1$ outperforms $\kest_{\mathrm{global}}$ is greater than $0.5$, that is that $\Prob_\mathbf{U}\left[J(\kest_1,\mathbf{U})<J(\kest_{\mathrm{global}},\mathbf{U})\right] > 0.5$.

The calibration of this numerical model is done in order to get a calibrated value $\kest$, that will be used for forecasts, using the same numerical model. We are now going to compare how well three of those values compare in a forecast context: $\kest_1$, $\hatkmean$ and $\kest_{\mathrm{global}}$, as the first two refer quite naturally to the notion of robustness, and the latter is an example of an estimate that does not take into account the uncertainties.
\clearpage
\subsection{Assessing the quality of the forecast of the calibrated model}
The calibration of the model has been performed using by integrating the model on a time-period $[0, T]$, called assimilation window. 
We now want to compare the quality of the different forecasts, from different calibrated bottom friction.
Those forecasts result from the integration of the numerical model between the time $T$ and a time $T_{\mathrm{pred}}$.

@@

\subsubsection{Squared forecast error}
Given two environmental conditions $\mathbf{u}$, used to run the computer simulation and $\mathbf{u}'$, used to generate the observations, the squared forecast error for the parameter $\mathbf{k}$ is
\begin{equation}
\label{eq:squared_forecast_error}
S_{\mathrm{pred}}(\mathbf{k},\mathbf{u},\mathbf{u}') = \left(\mathcal{M}_{\mathrm{pred}}(\mathbf{k},\mathbf{u}) - \mathcal{M}^{o}_{\mathrm{pred}}(\mathbf{u}')\right)^2
\end{equation}
Averaging over $\mathbf{u}$, $\mathbf{u}'$ defines the mean squared forecast error, defined on every point of the spatial domain, and at every time-steps. This can be done using a Monte-Carlo approximation: as $\mathbf{u}$ and $\mathbf{u}'$ are iid, let us assume that we have a set of samples $\{\mathbf{u}^i\}_{1\leq i \leq N_{\mathbf{u}}}$. 
\begin{equation}
  \label{eq:squared_forecast_error_averaged}
  S(\mathbf{k}) = \frac{1}{N_{\mathbf{u}}^2 }\sum_{i=1}^{N_{\mathbf{u}}}\sum_{j=1}^{N_{\mathbf{u}}} S_{\mathrm{pred}}(\mathbf{k},\mathbf{u}^i,\mathbf{u}^j) = \frac{1}{N_{\mathbf{u}}^2 }\sum_{i=1}^{N_{\mathbf{u}}}\sum_{j=1}^{N_{\mathbf{u}}}  \left(\mathcal{M}_{\mathrm{pred}}(\mathbf{k},\mathbf{u}^i) - \mathcal{M}^{o}_{\mathrm{pred}}(\mathbf{u}^j)\right)^2
\end{equation}
The mean squared forecast error averaged over the whole space and over all the time-steps gives an indication on the overall prediction quality of the prediction given this metric, and are represented on the right plot of Figure~\ref{fig:forecast_squared_error}. We can see that $\kest_{\mathrm{global}}$ performs slightly better than $\kest_1$, itself performing slightly better than $\hatkmean$.
Averaging $S(\mathbf{\kest}$ over only the time steps between $T$ and $T_{\mathrm{pred}}$, we have an indication on the quality of the forecast in the squared sense depending on the spatial position, as seen on the left plot of Figure~\ref{fig:forecast_squared_error}.
\begin{figure*}[!h]
\centering
\includegraphics[width=.9\textwidth]{Figures/forecast_error.pdf}
\caption{Quality of forecast, depending on the calibrated parameter. The left figures show the metric averaged over time, while the right figure show the mean forecast error, averaged over time and space}
\label{fig:forecast_squared_error}
\end{figure*}
We can see that the mean forecast error squared, on the right side, averaged over time and space is the smallest for $\kest_{\mathrm{global}}$, while $\kest_{1.0}$ performs slightly better $\hatkmean$.
\begin{figure}[!h]
  \centering
  \includegraphics[width=.9\textwidth]{Figures/forecast_CRPS}
  \caption{CRPS computed for different calibrated parameters}
\label{fig:forecast_crps}
\end{figure}

@@

\clearpage
\section*{Conclusion}  
\victor{À faire à la fin}
This paper deals with the problem of robust calibration of a computer code in the presence of uncertain inputs.
